---
title: "HPS302: Developmental Psychology (Adult)"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week 1: Introduction and research methods in adult development Pt 1  {.well .tabset .tabset-fade .tabset-pills}

### LO1

#### Contrast ‘individual differences’ to ‘commonalities’.

:::{.alert .alert-info}
- **Individual Differences** refer to the experiences and events in our lives that are individual to us, such as:
  - Single parenthood,
  - Grand parenthood,
  - Opting not to have children,
  - Living alone, 
  - Parents divorce.(e.g. Our family make-up, where we live, our neighbours, divorce, death, etc).
- **Commonalities** refer to the events and milestones in our lives that occur at roughly the same time, or are aspects of our lives that others can relate to such as:
  - Moving out of our parents’ homes,
  - Romantic relationships,
  - Entered university, and
  - Either started a family or given some serious thought to parenthood.
:::

:::{.alert .alert-info}
- **Stability**: The aspects of ourselves that are similar in the 20, 40 or 60 year old versions of ourselves. (i.e. interest in reading/books, interest in cycling)
- **Change**: The aspects of ourselves that are completely different to the younger or older versions of ourselves. (i.e. Developing an interest in travel, taking up a new hobby or interest, a new relationship, changing job).
  - **Outer**: Changes that are apparent to others such as losing/gaining weight, development of greater self-confidence, leaving interests and objects behind as we age.
  - **Inner**: Development of wisdom, from holding children close to giving them space, developing our own sense of family, etc.
:::

### LO2

#### Define and contrast the ideas of ‘continuous development’ and ‘staged development’.

:::{.alert .alert-info}
- **Continuous Development**: A slow and gradual development though the lifespan in learning and developing knowledge and hobbies such as gardening, cycling or renovating.
- **Staged Development**: Elements of the developmental journey that appear not to change for some time, but then change abruptly such as moving from single to married, spouse to parent, parent to grandparent, etc).
- Staged development can be broken down further into:
  - **Typical**: To what extent do we experiences change at the same time and in the same order? (i.e. Leaving high school to attend university)
  - **Atypical**: Are these life stages more individualised and unique to each of us? (i.e. Going from mother to student, or lecturer to student)
:::

### LO3

#### Explain each of the various sources of change and stability across adulthood.

:::{.alert .alert-info}
<h5><i class="fas fa-info-circle"></i> Sources of Change</h5>

- **Normative Age-Graded Influences**: the influences of age that affect most adults as they grow older:
  - **Biology**: A common part of aging referenced by our **biological clock**, such as greying hair, wrinkly skin, and losing of strength and muscle mass.
  - **Shared Experiences**: While not nearly as relevant for current generations as it was for previous generations, our **social clock** remains a guide for life events such as starting a family, leaving home, graduating from university, etc.
    - Our position on this social clock can impact our self-worth if we consider ourselves out of sync with societal expectations of the normative timing of life events.
    - Many consider these achievements outside the social norm to be considerable achievements.
    - Instances of **Ageism** are also governed by the social clock affecting both emerging adults and older adults.
    - Parenting can have a significant impact on a parent's social clock as experiences are governed by the requirements of a child's age and developmental stage, with parents being drawn together based on their shared experiences.
    - Experiences in this domain are far less likely to be universal than those governed by the biological clock.
  - **Internal Change Processes**: Parents have been observed to exemplify more traditional masculine and feminine traits as young adults, while older parents seek to balance or extend these traits.
  
</br>

- **Normative History-Graded Influences**: Historical events can shape both similarities and dissimilarities between people in similar groups (i.e. single child families in China)
  - **Culture** can shape the expected age of:
    - Marriage or childbearing,
    - The typical number of children (and wives),
    - The roles of men and women,
    - Class structures,
    - Religious practices, and
    - Laws.
  - **Cohort**: A group of people within a similar stage of life sharing similar historical experiences (i.e. Cuban Americans in the 60s, children of the Great Depression, children of COVID).
    - Smaller than a generation (SG, BB, X, Y, Z).
    - Can refer to the people within a country or region of a country.
    - The effects of the Great Depression differed between teens (showed greater independence and resilience) and primary school children (extended periods of hardship).
    - Other examples of the influences on cohorts include:
      - Economic conditions of the times,
      - Political and religious climate, 
      - Educational system, and 
      - Popular culture.
      
</br>      
      
- **Non-normative Life Events**
  - Examples of non-normative life events can include:
    - The death of a spouse in early adulthood,
    - Inheriting enough money to retire at age 40,
    - Taking over parental responsibility for grandchildren, and
    - Starting your own business at 65.
  - These events can be non-normative at any age (large inheritance), while others are non-normative due to the timing (death of a spouse).
  - The events that occur "on-time" can be easier to deal with than those that are "off-time".  

:::

:::{.alert .alert-info}
##### <i class="fas fa-info-circle"></i> Sources of Stability

- **Genetics (Nature)**
  - At conception, each individual inherits a unique combination of genes.
  - While most genes are shared across the human species (shaping common developmental patterns like walking at 12 months or puberty in the early teens), genetic inheritance also contributes to individual differences:  
    - The field of **behaviour genetics** studies how genes influence various traits, including:  
    - **Cognitive abilities** (e.g., IQ).  
    - **Physical traits** (e.g., height, body shape, metabolism).  
    - **Personality characteristics**.  
    - **Psychological disorders** (e.g., alcoholism, schizophrenia, depression).  
  - Research suggests that heredity plays a key role in **stability across development**, as many genetic traits and tendencies persist throughout life.  
  - Behaviour geneticists use **twin studies** to examine genetic influences on adult behaviour. These studies compare:  
    - **Monozygotic (identical) twins**, who share **100% of their genes**.  
    - **Dizygotic (fraternal) twins**, who share **about 50% of their genes**, like any other siblings.  
  - In a typical study, researchers measure a trait or ability in both twins and compare their similarity.
  - Research suggests that the higher similarity of identical than fraternal twins is more genetically influenced rather than shaped by environmental factors.  
  - Twin studies require **large sample sizes**, making recruitment difficult.\
  - To overcome this, countries with **centralised birth and health records** have taken the lead in this research.  
  - The **largest twin database** is at the **Karolinska Institute in Sweden**, which maintains records on **over 85,000 twin pairs**.  

</br>

- **Environment (Nurture)**
  - Both **genetics and environment** contribute to stability in personal development.
  - While neither determines destiny, both have lasting effects.  
  - The **Grant Study of Harvard Men**, led by psychiatrist **George Vaillant (2002)**, found that:  
    - Men from **warm, trusting homes** were more likely to lead **well-adjusted** adult lives.  
    - They expressed emotions openly, trusted others, and maintained strong friendships.  
    - Parental warmth fostered **self-worth, coping skills, and meaningful relationships**.  
  - Vaillant’s research also linked **early childhood environment** to **successful ageing**:  
    - Those from supportive homes were **healthier and happier** at age 75.  
    - Those from bleak homes faced **loneliness, mistrust, and illness** later in life.  
    - These findings contributed to **Vaillant’s theory of personality development**.  

</br>

- **Interactionist View**
  - There is no clear division between **genes and environment**, as they interact to shape stability throughout adulthood.
  - The **interactionist perspective** suggests that genetic traits influence how individuals engage with their environment, which in turn reinforces those traits.  
  - For example:  
    - A **risk-averse child** may seek low-risk activities, form friendships with similar individuals, and be guided toward a cautious career path.  
    - Over time, their lifestyle continues to reflect **risk avoidance**, reinforcing their behaviours and experiences.  
  - The interactionist view sees Nature vs. Nurture as a **chicken-and-egg dilemma**.  
  - A biological mechanism, **epigenetic inheritance**, explains how **environmental experiences modify gene expression**.
  - This process, called **DNA methylation**, alters genes by adding a methyl group, reducing gene expression.  
  - **Example**
    - **Childhood Abuse and Lifelong Genetic Changes**  
      - Research on **suicide victims** found that those with a history of childhood abuse had modified **glucocorticoid receptor genes**.
      - This influences **stress responses** compared to those without a history of childhood abuse.  
      - This suggests that **early life experiences** can lead to **permanent genetic changes** with lifelong consequences.  
  - These findings highlight the complex interplay between **genetics and environment**, shaping development in ways far beyond what is inherited at birth.b

:::

### LO4

#### Explain the difference between normative age-graded influences, normative history-graded influences and non normative life events and the potential impact of each on adult development

::: {.alert .alert-warning}
<h5><i class="fas fa-info-circle"></i> On Ageing</h5>

- **Chronological Age and Development**  
  - **Chronological age** refers to the number of years since birth, often marked by birthdays.
  - While it is significant in **childhood** (as children of the same age tend to have similar abilities and interests), it becomes **less relevant in adulthood**, except in specific cases like:  
    - **Young adulthood** – Legal milestones (e.g., driving, purchasing alcohol, voting).  
    - **Older adulthood** – Eligibility for benefits (e.g., Social Security, Medicare).  
  - However, **chronological age does not cause development**—aging is not simply a result of time passing or birthday candles burning. While **age may be related to developmental changes, it is not the direct cause**.  

</br>

- **Biological Age and Physical Condition**  
  - **Biological age** measures how an individual’s **physical condition** compares to others of different ages. Informal examples include:  
    - “He has the **memory of a 50-year-old**.”  
    - “She **runs like a 30-year-old**.”
  - The meaning of **biological age** depends on **chronological age** (a 30-year-old with the memory of a 50-year-old may raise concerns), whereas a 70-year-old with the same memory would seem youthful.  
  - **Biological age is often assessed through medical evaluations**, such as **bone density scans**, comparing patients' bones to those of a healthy 20-year-old.
  - **Lifestyle changes** can influence biological age, potentially slowing or reversing some aspects of aging.  
  
</br>

- **Psychological, Social, and Functional Age**  
  - **Psychological age** refers to an individual’s ability to effectively navigate life compared to others. For example:  
    - A **30-year-old** who overspends and is often late for work functions more like a **teenager**.  
- **Social age** is based on the roles a person assumes at different life stages:  
  - A **40-year-old first-time mother** has a **social age** younger than her chronological age.  
  - A **23-year-old supporting her family** has a **social age** beyond her years.  

</br>

- **Functional Age and Measuring Development**  
  - Sometimes, **biological, psychological, and social age** are grouped as **functional age**, reflecting **overall adult functioning**.  
  - Developmental psychologists recognise that **chronological age alone is not a reliable measure** of adult behaviour. Instead, studies often compare:  
    - **Different age groups** (e.g., young vs. middle-aged adults).  
    - **Different roles** (e.g., couples with vs. without children).  
    - **The same individuals before and after major life transitions** (e.g., parenthood, retirement).  
  - This highlights that **development and chronological age do not always align**, especially as people grow older.  

:::

### LO5

#### Describe Baltes’ Lifespan Approach (including Table 1.2 on p. 34) and Bronfenbrenner’s Bioecological Model. Describe how each approach would explain an individual’s development, and consider any similarities between these two models.

:::{.alert .alert-info}
<h5><i class="fas fa-info-circle"></i> Life-Span Developmental Psychology Approach</h5>

- The **life-span developmental psychology approach** views development as **lifelong, multidimensional, plastic, contextual, and influenced by multiple causes**.  
- Introduced by **Paul Baltes et al. (1980)**, this approach marked a turning point in the field, which had previously focused almost entirely on **child development**.
- It expanded the study of development to **all life stages**, from childhood to old age.  
- By recognising development as a **continuous and complex process**, this perspective allows researchers to study not just children but also **adults of all ages**, including **students, parents, professors, and grandparents**.  
- This framework has become fundamental in modern developmental psychology, shaping how we understand **growth and change across the entire lifespan**. 

</br>

<table class="table table-striped table-bordered">
<thead>
<tr>
<th>Concept</th>
<th>Proposition</th>
<th>Example</th>
</tr>
</thead>
<tbody>
<tr>
<td>Life-span development</td>
<td>Human development is a lifelong process. No single age is more important than another.</td>
<td>A 38-year-old single woman makes plans to adopt a child; a 75-year-old Civil War buff begins taking a class in memoir writing.</td>
</tr>
<tr>
<td>Multidirectionality</td>
<td>We develop in different directions and at different rates. Some abilities increase while others decline.</td>
<td>Young adults become independent when starting a career but remain dependent when living with parents.</td>
</tr>
<tr>
<td>Development as gain and loss</td>
<td>Development is a mix of gains and losses at all ages, requiring adaptation.</td>
<td>Middle-aged adults may lose their parents but gain a new sense of maturity.</td>
</tr>
<tr>
<td>Plasticity</td>
<td>Many aspects of development can change, but there are limits.</td>
<td>Young adults with behaviour problems can become responsible; fathers can stay home with kids while mothers work.</td>
</tr>
<tr>
<td>Historical embeddedness</td>
<td>Development is influenced by historical and cultural conditions.</td>
<td>People who grew up in the 1970s have more open attitudes toward legalising drugs than earlier or later cohorts.</td>
</tr>
<tr>
<td>Contextualism</td>
<td>Development is shaped by age-graded, history-graded, and non-normative influences.</td>
<td>Each of us is unique due to shared generational experiences and personal life events.</td>
</tr>
<tr>
<td>Multidisciplinary</td>
<td>Human development is studied across multiple disciplines, not just psychology.</td>
<td>Fields like sociology, economics, public health, and education contribute to understanding development.</td>
</tr>
</tbody>
</table>
:::

:::{.alert .alert-info}
<h5><i class="fas fa-info-circle"></i> Bioecological Model of Development</h5>

- The **bioecological model** emphasizes that human development occurs within multiple interacting environments—biological, psychological, and social—which change over time.
- Introduced by **Urie Bronfenbrenner** in 1979 and later refined, the model consists of five systems:
  - **Microsystem** (immediate environment like family and friends)  
  - **Mesosystem** (interactions between elements in the microsystem)  
  - **Exosystem** (indirect influences, such as parental workplaces)  
  - **Macrosystem** (cultural and societal influences)  
  - **Chronosystem** (changes over time, including personal growth and historical events)  
- The key idea is that **development cannot be studied in isolation**—it must be understood in the context of relationships, communities, and broader cultural influences.
- Modern research increasingly applies this model to study adults in various roles, including as individuals, partners, parents, workers, and members of cultural groups.

:::

### LO6

#### Explain the major tasks (or goals) of each stage of adulthood

:::{.alert .alert-info}
<h5><i class="fas fa-info-circle"></i> Major goals of adulthood</h5>

<table class="table table-bordered table-striped">
  <thead>
    <tr>
      <th>Life Stage</th>
      <th>Major Tasks</th>
    </tr>
  </thead>
  <tbody>
    <tr>
      <td>Emerging Adulthood (18–24 Years)</td>
      <td>Establish self as an adult by completing education or job training, become independent from parents financially, make own decisions, become an autonomous member of the community.</td>
    </tr>
    <tr>
      <td>Early Adulthood (25–39 Years)</td>
      <td>Renegotiate relationship with parents; form intimate partnerships; begin family; begin career; create individual identity; strive for success in both personal and professional life.</td>
    </tr>
    <tr>
      <td>Middle Adulthood (40–64 Years)</td>
      <td>Guide children into adulthood; cope with death of parents; strengthen marriage; redefine life goals; achieve individuality; care for aging family members.</td>
    </tr>
    <tr>
      <td>Older Adulthood (65–74 Years)</td>
      <td>Find an alternative to a lifelong job; cope with health problems of self and spouse; redefine life goals and sense of self.</td>
    </tr>
    <tr>
      <td>Late Adulthood (75+ Years)</td>
      <td>Come to terms with one’s own life, possibly through reviewing memories or writing a memoir. Cope with the deaths of loved ones and the eventuality of own death. Value remaining family members, friends, and other joys in life.</td>
    </tr>
  </tbody>
</table>

:::

### LO7

#### Explain the changes in the domains of adult functioning (physical, cognitive, social, work and personality) across each stage of adulthood 

:::{.alert .alert-info}

```{r, echo=FALSE, results='asis'}
library(knitr)
library(kableExtra)

data <- data.frame(
  Category = c(
    "Physical Change", "Cognitive Change", "Family and Gender Roles",
    "Relationships", "Work Roles", "Personality and Meaning"
  ),
  `Emerging Adulthood (18–24 Years)` = c(
    "Peak functioning in most body systems and physical abilities; optimal reproductive years; health habits established now will create pattern of later well-being. Bone mass is still increasing.",
    "Peak period of cognitive skill on most measures, fastest reaction time. Most depend on parents for important decisions. Driving safety is low, and electronics use is high and skilled.",
    "Family roles are mixture of childhood and adult roles, and young people move in and out of them.",
    "Family relationships are similar to childhood; peers are important socially. Romantic relationships consist of short-term dating for most.",
    "Vocational interests present for most. Jobs are part-time or entry-level and not related to vocational interests. Job performance is variable. Little thought of retirement plans. Residential moves are for college or military service.",
    "Establish identity in areas of occupation, gender role, political, and religious beliefs. Begin to see the self as separate from the group. Accentuate own gender characteristics and repress opposite-gender characteristics."
  ),
  `Early Adulthood (25–39 Years)` = c(
    "Bone mass is still increasing. Obesity is present for some, and many do not follow diet and exercise recommendations.",
    "Most memory abilities show slight but gradual declines except crystallized abilities, semantic and procedural memory. Decision making increases. Driving safety increases. Electronic use is an integral part of lives.",
    "Major family roles are acquired (e.g., spouse, parent). Advances in these areas dominate life. Clear separation of gender roles.",
    "Emphasis on forming new friendships, cohabitation, and marriage. Continued relationship with parents, siblings, and often grandparents.",
    "Emphasis on choosing career, changing jobs, and establishing oneself in a career.",
    "Increased individuality (self-confidence, independence, autonomy). Form own ideas and standards. Gender differences remain high."
  ),
  `Middle Adulthood (40–64 Years)` = c(
    "High levels of functioning continue (except for top athletes in most sports). Slight decline in sensory functioning appears. Weight and girth increase. Bone mass begins slight decline.",
    "Small declines continue for all except crystallized abilities (which peak), semantic and procedural memory. Decision making increases due to experience. Driving safety is good. Electronic use is variable.",
    "Launch children; postparental phase; for many, added role of caregiver for elderly family members. Grandparent role begins for most.",
    "Increased marital satisfaction as focus turns from parenting to other relationships. Adult children remain important.",
    "Peak years of career success and income for most, also work satisfaction. Slight physical and cognitive decline is compensated for by increases in job expertise.",
    "Establish generativity within family or workplace. Some sign of a softening of the individuality of the earlier period; fewer immature defenses; possibly autonomous level. Increase in spirituality for some, especially women."
  ),
  `Older Adulthood (65–74 Years)` = c(
    "Physical decline more noticeable, but rate of decline is still relatively slow; reaction time slows. Weight decreases for many as bone and muscle mass decrease. More decline in sensory systems.",
    "More slow decline for memory systems except crystallized abilities and procedural memory. Decision making remains good. Driving begins to decline, but may be helped with training. Electronic use can be valuable to help with day-to-day tasks.",
    "Grandparent role continues in importance; significantly less dominance of gender roles.",
    "High marital satisfaction for those who have spouses; friendships and sibling relationships may become more intimate. Relationships with adult children frequent but not central to well-being. Relationship with grandchildren is important.",
    "Most leave their full-time jobs and take less stressful or part-time jobs, do volunteer work, or retire entirely.",
    "Task of ego integrity; perhaps more interiority; a few may reach integrated level. Increase in spirituality for most."
  ),
  `Late Adulthood (75+ Years)` = c(
    "Acceleration in decline of physical abilities and sensory functioning.",
    "All systems have gradual decline, including crystallized abilities, but not procedural memory. Decision making can be affected by cognitive loss and health disorders (and prescription drugs). Driving safety declines sharply. Electronic use continues to be helpful if started at earlier age.",
    "Participation in family roles declines as activities are restricted. Role of care receiver begins for some.",
    "Majority are widowed; small network of close friends and siblings remain important.",
    "Work roles unimportant for most. Some continue to volunteer.",
    "Continuation of previous pattern. Increase in spirituality for most, even if outward signs decrease."
  ),
  check.names = FALSE
)

kable(data, format = "html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  column_spec(1, bold = TRUE)
```

:::

### LO8

#### Explain the purpose of thematic analysis

:::{.alert .alert-info}

Thematic analysis is a **qualitative research** method used to identify, analyse, and interpret patterns (themes) within textual data. It helps researchers make sense of large amounts of information by categorizing recurring ideas, meanings, or experiences.

- **Key Steps in Thematic Analysis:**
  1. **Familiarization** – Reading and re-reading the data to gain an overall understanding.
  2. **Coding** – Identifying meaningful segments of text and assigning labels (codes) to them.
  3. **Searching for Themes** – Grouping similar codes into broader themes that represent key patterns in the data.
  4. **Reviewing Themes** – Refining themes by checking their relevance and coherence with the dataset.
  5. **Defining and Naming Themes** – Clearly articulating what each theme represents.
  6. **Writing the Report** – Presenting findings with supporting examples from the data.

</br>

- **Types of Thematic Analysis:**
  - **Inductive (Bottom-Up):** Themes emerge directly from the data without predefined categories.
  - **Deductive (Top-Down):** Themes are based on pre-existing theories or frameworks.
  - **Semantic Analysis:** Focuses on explicit meanings in the data.
  - **Latent Analysis:** Examines underlying meanings and assumptions behind the words.

:::

### LO9

#### Explain the meaning of reflexivity and what would be required for a researcher to become a 'reflexive researcher'

:::{.alert .alert-info}

- **Meaning of Reflexivity**  
  - Reflexivity in research refers to the process of critically examining one’s own role, biases, and influence on the research process.
  - It involves acknowledging how personal beliefs, experiences, and perspectives shape the research design, data collection, analysis, and interpretation.
  - Reflexivity helps ensure that research findings are not unduly influenced by the researcher’s subjectivity but are instead thoughtfully examined in relation to it.

</br>

- **Becoming a Reflexive Researcher**: To be a reflexive researcher, one must engage in continuous self-awareness and critical reflection throughout the research process. This includes:
  1. **Acknowledging Personal Biases & Assumptions**  
    - Recognizing one’s own values, cultural background, and preconceptions that may influence the study.
    - Reflecting on how these factors shape interactions with participants and the interpretation of data.
  2. **Engaging in Reflexive Writing**  
    - Keeping a research journal or memo to document thoughts, feelings, and potential biases.
    - Writing reflexive statements in research papers to acknowledge the researcher’s positionality.
  3. **Being Transparent about Researcher-Participant Dynamics**  
    - Considering how power relationships, identity, and social position affect the research process.
    - Being aware of how participants’ responses might be influenced by the researcher’s presence.
  4. **Using Multiple Perspectives & Methods**  
    - Triangulating data sources to minimize bias.
    - Seeking feedback from peers or participants to validate findings.
  5. **Continuously Revisiting Assumptions**  
    - Regularly questioning one's interpretations and being open to alternative explanations.
    - Engaging in discussions with colleagues or mentors to challenge personal viewpoints.

</br>

- **Why Reflexivity is Important**  
  - Enhances the credibility and trustworthiness of qualitative research.  
  - Helps researchers produce more ethical and authentic findings.  
  - Allows for deeper, more nuanced interpretations of data.  

:::

### LO10

#### Explain the key differences between a qualitative and quantitative orientation   

:::{.alert .alert-info}

```{r, echo=FALSE, results='asis'}
library(knitr)
library(kableExtra)

# Data for qualitative vs quantitative comparison
data <- data.frame(
  Aspect = c(
    "Purpose", "Nature of Data", "Data Collection Methods", "Analysis Approach", 
    "Sample Size", "Research Design", "Findings", "Researcher’s Role", 
    "Types of Questions", "Outcome"
  ),
  `Qualitative Orientation` = c(
    "Explores meanings, experiences, and social phenomena in depth.",
    "Non-numerical (words, images, observations).",
    "Interviews, focus groups, observations, open-ended surveys, case studies.",
    "Thematic analysis, discourse analysis, grounded theory, content analysis.",
    "Small, purposive, or theoretical sampling (not focused on generalizability).",
    "Flexible, emergent, and exploratory.",
    "Rich, detailed, and subjective, often presented in narratives.",
    "Actively engaged in data interpretation; reflexivity is important.",
    "Open-ended, exploratory (e.g., 'How do people experience stress?').",
    "Develops deep understanding, theories, and insights."
  ),
  `Quantitative Orientation` = c(
    "Measures variables and analyzes numerical data to identify patterns and relationships.",
    "Numerical (numbers, statistics, measurements).",
    "Surveys, experiments, structured observations, existing datasets.",
    "Statistical analysis (descriptive, inferential, correlations, regressions).",
    "Large, random, or representative samples for generalizability.",
    "Structured, standardized, and hypothesis-driven.",
    "Objective, statistical, and generalizable to larger populations.",
    "Maintains distance to minimize bias; objectivity is key.",
    "Closed-ended, measurable (e.g., 'What percentage of people experience stress?').",
    "Tests hypotheses, predicts trends, and establishes causal relationships."
  ),
  check.names = FALSE
)

# Generate HTML table with Bootstrap 3.3.5 styling
kable(data, format = "html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  column_spec(1, bold = TRUE)
```

- **Choosing Between the Two Approaches**
  - Use Qualitative Research when exploring subjective experiences, meanings, and complex social phenomena.
  - Use Quantitative Research when measuring variables, testing hypotheses, or seeking statistical generalizations.
  - Mixed methods combine both approaches to provide a more comprehensive understanding.
:::

## Week 2: Research methods in adult development Pt 2: Thematic analysis {.well .tabset .tabset-fade .tabset-pills}

- Understand the meaning of qualitative research (Lecture)
  - Identify the methods associated with qualitative research
  - Explain the paradigms associated with qualitative research and how they differ from the paradigms associated with quantitative research
- Understand the meaning and purpose of reflexive thematic analysis (Lecture)
  - Describe the capabilities and limitations of RTA as a method
  - Explain what it means to be a reflexive researcher and how this is achieved
- Understand the importance of RTA to developmental psychology (Lecture)
  - Compare research questions which are answered with RTA to questions answerable using quantitative research approaches
- Describe the stages of RTA (Lecture)
- Familiarise yourself with data (Lecture)
- Conduct coding (pp. 53 - 75) 
- Create and refine themes (pp. 76 - 111)
- Write up a RTA (Lecture)

### LO1

:::{.alert .alert-info}
#### <i class="fas fa-info-circle"></i> Understand the meaning of qualitative research (Lecture)

- RTA is a form of qualitative research
  - Qualitative research seeks to collect, analyse and report on word data, through:
    - Interviews
    - Open ended survey questions
    - Online posts

>"&hellip;qualitative research involves an interpretive, naturalistic approach to the world. This means that qualitative researchers study things in their natural settings, attempting to make sense of, or to interpret, phenomena in terms of the meanings people bring to them"
>
><footer class="pull-right">Denzin and Lincoln (2005)</footer>

- Qualitative research explores human experiences, meanings, and social phenomena through word-based data.
- Unlike quantitative research (which focuses on numerical measurement), qualitative research seeks to understand how people make sense of their worlds by prioritising depth, complexity, and context, through:
  - interviews,
  - focus groups, and
  - textual analysis to understand how people make sense of their worlds.
- Qualitative research refers to more than just the methods we use or the type of data that we collect.
- It also refers to a philosophical view (often called **paradigms**) that reality (**ontology**) is not objective and that knowledge (**epistemology**) is shaped by context, researcher interpretation and previous experiences.
- Braun and Clarke (2021) consider qualitative research based in this philosophy as a genuine form of qualitative research and label this approach **'Big Q'**.
- Approaches which do not follow the 'Big Q' view of qualitative research are called **'Small Q'**.
- These approaches apply a positivist or post-positivist paradigm and treat reality as objective and not highly influenced by experiences, power dynamics or social context. 
- Reflexive Thematic Analysis is aligned with a 'Big Q' approach to qualitative research.
:::

:::{.alert .alert-info}
#### <i class="fas fa-info-circle"></i>Identify the methods associated with qualitative research

:::
:::{.alert .alert-info}
#### <i class="fas fa-info-circle"></i>Explain the paradigms associated with qualitative research and how they differ from the paradigms associated with quantitative research

:::

### LO2

:::{.alert .alert-info}
#### <i class="fas fa-info-circle"></i> Understand the meaning and purpose of reflexive thematic analysis (Lecture)

:::

#### <i class="fas fa-info-circle"></i> Describe the capabilities and limitations of RTA as a method

#### <i class="fas fa-info-circle"></i> Explain what it means to be a reflexive researcher and how this is achieved

### LO3

Understand the importance of RTA to developmental psychology (Lecture)

#### Compare research questions which are answered with RTA to questions answerable using quantitative research approaches

### LO4

:::{.alert .alert-info}
#### <i class="fas fa-info-circle"></i> Describe the stages of RTA (Lecture)  

1. **Familiarisation with Data** – Immerse yourself in the dataset by reading, re-reading, and making notes on key insights. If working with audio, listen to recordings at least once.  
2. **Coding** – Systematically identify and label meaningful data segments. Codes capture **explicit** (semantic) or **implicit** (latent) meanings and reflect the researcher’s analytic perspective.  
3. **Generating Initial Themes** – Cluster related codes to construct broader themes that help answer the research question. Themes are actively developed rather than simply "discovered."  
4. **Developing & Reviewing Themes** – Assess whether themes accurately reflect the data. This may involve merging, splitting, or discarding themes to ensure coherence and relevance.  
5. **Refining & Naming Themes** – Fine-tune each theme, ensuring clarity and a strong core concept. Assign clear, concise names and describe how they contribute to the overall analysis.  
6. **Writing Up** – Integrate themes into a compelling analytic narrative, using vivid data extracts. The final report includes an introduction, method, analysis, and conclusion, with thorough editing.
:::

### LO5

Familiarise yourself with data (Lecture)

#### Topic 5

### LO6

Conduct coding (pp. 53 - 75)

####  Topic 6

### LO7

Create and refine themes (pp. 76 - 111)

#### Topic 7

### LO8

Write up a RTA (Lecture)

#### Topic 8

## Week 3: Social roles in adulthood {.well .tabset .tabset-fade .tabset-pills}

### LO1

#### Topic 1

### LO2

#### Topic 2

### LO3

#### Topic 3

### LO4

#### Topic 4

### LO5

#### Topic 5

### LO6

#### Topic 6

### LO7

#### Topic 7

## Week 4: Social relationships in adulthood {.well .tabset .tabset-fade .tabset-pills}

### LO1

#### Topic 1

### LO2

#### Topic 2

### LO3

#### Topic 3

### LO4

#### Topic 4

### LO5

#### Topic 5

### LO6

#### Topic 6

### LO7

#### Topic 7

## Week 5: Personality continuity and change over adulthood {.well .tabset .tabset-fade .tabset-pills}

### LO1

#### Topic 1

### LO2

#### Topic 2

### LO3

#### Topic 3

### LO4

#### Topic 4

### LO5

#### Topic 5

### LO6

#### Topic 6

### LO7

#### Topic 7

## Week 6: Physical changes across adulthood {.well .tabset .tabset-fade .tabset-pills}

### LO1

#### Topic 1

### LO2

#### Topic 2

### LO3

#### Topic 3

### LO4

#### Topic 4

### LO5

#### Topic 5

### LO6

#### Topic 6

### LO7

#### Topic 7

## Week 7: Cognitive changes across adulthood Pt 1 {.well .tabset .tabset-fade .tabset-pills}

### LO1

#### Topic 1

### LO2

#### Topic 2

### LO3

#### Topic 3

### LO4

#### Topic 4

### LO5

#### Topic 5

### LO6

#### Topic 6

### LO7

#### Topic 7

## Week 8: Cognitive changes across adulthood Pt 2: Neurocognitive disorders {.well .tabset .tabset-fade .tabset-pills}

### LO1

#### Topic 1

### LO2

#### Topic 2

### LO3

#### Topic 3

### LO4

#### Topic 4

### LO5

#### Topic 5

### LO6

#### Topic 6

### LO7

#### Topic 7

#### Motor Imagery and Motor Skill Interventions
## Week 9: Long-term care and ageing in place {.well .tabset .tabset-fade .tabset-pills}

### LO1

#### Topic 1

### LO2

#### Topic 2

### LO3

#### Topic 3

### LO4

#### Topic 4

### LO5

#### Topic 5

### LO6

#### Topic 6

### LO7

#### Topic 7

## Week 10: Stress and resilience {.well .tabset .tabset-fade .tabset-pills}

### LO1

#### Topic 1

### LO2

#### Topic 2

### LO3

#### Topic 3

### LO4

#### Topic 4

### LO5

#### Topic 5

### LO6

#### Topic 6

### LO7

#### Topic 7
## Week 11: Preparing for the end of life {.well .tabset .tabset-fade .tabset-pills}

### LO1

#### Topic 1

### LO2

#### Topic 2

### LO3

#### Topic 3

### LO4

#### Topic 4

### LO5

#### Topic 5

### LO6

#### Topic 6

### LO7

#### Topic 7