---
title: "HPY210: Applied Counselling Skills (Introductory)"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(kableExtra)  # For additional styling (optional)
library(ggplot2)
```

## Week 1: Introduction and Research Methods in Adult Development Pt 1  {.well .tabset .tabset-fade .tabset-pills}

### Lecture 1: Introduction to Counselling

:::{.alert .alert-info}

#### <i class="fas fa-info-circle"></i> How is Counselling Defined?

</br>

##### Professional Organisations

::: {.blockquote}
>“Professional counselling is a safe and confidential collaboration between **qualified** counsellors and clients to promote **mental health and wellbeing, enhance self-understanding**, and resolve identified concerns. **Clients are active participants** in the counselling process at every stage”
>
><footer class="pull-right">Psychotherapy and Counselling Federation of Australia (PACFA)</footer>
:::

:::{.blockquote}
>Counselling is a professional **relationship** that **empowers** diverse individuals, families, and groups to accomplish **mental health, wellness, education, and career goals**.
>
><footer class="pull-right">American Counselling Association</footer>
:::

:::{.blockquote}
>Counselling and psychotherapy are **umbrella terms** that cover a range of talking therapies. They are delivered by **trained practitioners** who work with people over a short or long term to help them bring about **effective change or enhance their wellbeing**.
>
><footer class="pull-right">British Association for Counselling and Psychotherapy</footer>
:::

</br>

##### Textbooks

::: {.blockquote}
>"&hellip;a **principled relationship** characterised by the application of one or more psychological **theories and a recognised set of communication skills**, modified by experience, intuition and other interpersonal factors, to **clients’ intimate concerns, problems or aspirations.**”
>
><footer class="pull-right">Feltham and Dryden (1993)</footer>
:::

::: {.blockquote}
>“Counselling is an **intimate form of learning**, and it demands a practitioner who is willing to be an authentic person in the **therapeutic relationship.** It is within the context of such a **person-to-person connection** that the client experiences **growth.**”
>
><footer class="pull-right">Corey (2016)</footer>
:::

::: {.blockquote}
>“Counselling is [an] **intensive and personal process**&hellip;[it is] about **listening to and understanding** a client’s life challenges and **developing strategies** for change and growth.”
>
><footer class="pull-right">Ivey, Ivey and Zalaquett (2018)</footer>
:::

</br>

##### Key Components of Effective Counselling

| **Concept**                          | **Description** |
|--------------------------------------|--------------------------------------------------------------------------------------------------------------------------------|
| **Relationship**                     | The relationship is built on trust, empathy, privacy and respect, and it allows the client to feel heard and understood.       |
| **Client-centredness**               | Counselling is a purposeful endeavour that arises from and serves the client’s intentions, priorities, needs, hopes and goals. |
| **Understanding and Change**         | Counselling stimulates capacity for understanding and fosters learning of skills/strategies for pursuit of personal goals or mental wellbeing. |
| **Skilful (and qualified) use of techniques** | Effective counselling relies on the mastery of a set of skills and approaches. Accreditation organisations monitor competencies for professional counsellors. |

</br>

##### Key Ingredients of Effective Counselling

![](images/HPY210/Week1/Effective.svg)

##### Professional Counselling vs. Counselling Skills

- Professional Counselling is  :
  - Conducted by Psychologists, counsellors, psychotherapists, social workers, psychiatrists.
  - Uses a therapeutic approach to promote personal development or enhance mental wellbeing.
  - Delivered over a number of sessions (short-term or long-term), each session has limited duration (usually 50 mins).
- Counselling Skills are:
  - Used in a variety of careers, such as people-focused healthcare professionals, coaches, trainers/facilitators, supervisors, managers, tutors and presenters.
  - Aims at enhancing personal development within a particular domain (current employment or group).
  - Delivered more casually on an ad-hoc basis.

##### Counselling Skills vs. Communication/Interpersonal Skills

- Counselling Skills involve:
  - Attentive listening
  - Responding to and focusing on the client’s interests.
  - Confidentiality
  - Application of microskills
  - Goal-focused

##### Comparison Between Counselling and Coaching

- **Counselling**:  
  - Awareness.
  - Assistance with problems or dysfunction.
  - Getting back to ‘normal’ functioning.
  - Has tended until recently to be problem focussed.
- **Coaching**:
  - Awareness;
  - Skill development;
  - Going beyond current capabilities;
  - Building strengths; Often outcome-focussed;
  - Business and sport focus, and within training organisations;
  - Health
- **Similarities** identified (between coaching & counselling):
  - Listening
  - Questioning
  - Non-judgemental (a ‘safe’ space)
  - The process of uncovering (although what was uncovered differed)
- **Differences** identified:
  - Focus and intention (coaching aimed at movement toward desired outcomes; present-future orientated; realising potential)
  - Client base (coaching aimed at business, but can venture beyond).
  
::: {.blockquote}
>“Coaching [is] a human development process that involves structured, focused interaction and the use of appropriate strategies, tools and techniques to promote desirable and sustainable change&hellip;”  
><footer class="pull-right">Cox et al, (2014)</footer>  
:::

##### How Can Counselling Help?

- **Insight**: gaining understanding of the origins and development of one’s difficulties and <abbr title="enduring set of behaviours reflecting a pervasive/inflexible response-style, which may at one stage have helped with self-preservation, it but has now become deleterious to well-being.">self-defeating patterns</abbr>.
- **Awareness**: the process of noticing and experiencing one’s thoughts, feelings, and behaviours in real time.
- **Self-acceptance**: development of a positive attitude towards self.
- **Relating with others**: ability to form and maintain satisfying relationships with others.
- **Problem-solving**: a cognitive and behavioural process of finding solutions to problems and includes a series of steps.
- **Psychoeducation**: a collaborative process of information-giving and skills-building to promote changes in thoughts, feelings and behaviours.
- **Cognitive change**: a process of modifying or restructuring the thought patterns and core beliefs underlying psychological distress.
- **Behaviour change**: empowering individuals to adopt positive practices
- **Distress reduction**
- **tolerance and emotion regulation**

</br>

:::{.blockquote}
> “Feelings, emotions, and moods constitute a river that continually runs through us – peaceful, meandering, turbulent, or raging, often beneficial, sometimes dangerousm seldom neutral”.
>
><footer class="pull-right">Egan (2014)</footer>
:::

</br>

:::{.blockquote}
> “To change in therapy, clients cannot just talk intellectually about themselves and their feelings, they need to viscerally experience what they talk about&hellip;”
>
><footer class="pull-right">Greenberg (2002)</footer>
:::

</br>

##### Working With Feelings

- According to Greenberg (2010) there are three categories or strategies when working with feelings:
  - **Building awareness through research**; putting feelings into words helps decrease amygdala arousal (Lieberman et al 2004).
  - **Expressing feelings**: Feelings are elicited and tolerated in the session; using body awareness to help clients get in touch with their feelings.
  - **Reflection and regulation**.

:::

### Lecture 2: When and Why Do People Change?

### Compulsory Reading

```{r, echo=FALSE, results='asis'}

# Data for counselling process table
data <- data.frame(
  `Process of Counselling` = c(
    "Developing trust and engagement, including ‘presence’ and empowerment through relationship change",
    "Exploring the problem",
    "Goal-setting",
    "Empowering action",
    "Maintaining the change",
    "Ending counselling"
  ),
  `Cycle of Change` = c(
    "Pre-contemplation",
    "Contemplation",
    "Preparation",
    "Action",
    "Maintenance",
    "Termination or lapse or relapse"
  ),
  `Process of Maintaining Change` = c(
    "Anticipating and preventing problems",
    "Stopping the triggering of a return to problematic behaviour",
    "Monitoring and analysing high-risk situations",
    "Teaching coping responses",
    "Stopping a lapse become a relapse",
    "Dealing with helplessness; reducing expectations of old patterns of thinking and behaving"
  ),
  check.names = FALSE
)

# Generate HTML table with Bootstrap 3.3.5 styling
kable(data, format = "html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  column_spec(1, bold = TRUE)

```



## Week 2: Introduction and Research Methods in Adult Development Pt 2  {.well .tabset .tabset-fade .tabset-pills}

## Week 3: Ethics and Reflective Practice  {.well .tabset .tabset-fade .tabset-pills}

## Week 4: Ethical Decision Making Frameworks and Risk Assessment  {.well .tabset .tabset-fade .tabset-pills}

## Week 5: Theories and Frameworks: Introduction, Assessments, and Terminations  {.well .tabset .tabset-fade .tabset-pills}

## Week 6: Theories and Frameworks: Person-Centred Approaches  {.well .tabset .tabset-fade .tabset-pills}

## Week 7: Theories and Frameworks: Motivational Interviewing  {.well .tabset .tabset-fade .tabset-pills}

## Week 8: Theories and Frameworks - Solution Focused Approaches  {.well .tabset .tabset-fade .tabset-pills}

## Week 9: Theories and Frameworks - Cognitive Behavioural Approaches  {.well .tabset .tabset-fade .tabset-pills}

## Week 10: Theories and Frameworks - Mindfulness Based Approaches  {.well .tabset .tabset-fade .tabset-pills}

